Conversations in time: interactive visualisation to explore structured temporal data

### 2020-09-06 tsibbletalk.Rmd
explained wrapping and shiny module functions

The city of Melbourne has sensors installed to count hourly tallies of pedestrians in order to capture downtown daily rhythms (cite). Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four locations, with the depiction of three pronounced slices in time. Figure \ref{fig:wrap-ped-1} unfolds all counts from January to May on their absolute time scales, facetted by four sensors. On March 16, Melbourne went to the stage three lockdown due to COVID-19, seeing a significant decline in traffic volume across the city. These lines are then folded into daily and weekly sections, shown as Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} respectively. Seasonal variations have been popped out to viewers, complementing the not-just-magnitude-drop data story. The pre-lockdown period is coloured with dark green and lockdown with orange.

Figure \ref{fig:wrap-ped} suggests that there could be more than one interesting time slices. The \CRANpkg{tsibbletalk} package provides a pair of UI and server functions, as a shiny module, to help with finding interesting seasonal slices in a shiny application. The pair, decoupled to `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece. Like all shiny modules, users should supply a unique session id. The UI function `tsibbleDiceUI()` shows a slider that controls the number of periods, and a plot specified by users. The server function `tsibbleDiceServer()` is the workhorse, transforming data and updating the plot. It expects a `ggplot` (converted to `plotly` via `ggplotly()`) or `plotly` object. This plot can be line charts, or other graphical elements (such as boxplots). But it assumes that tsibble's time index is plotted on the x-axis. The other mandatory argument is to specify the number of seasonal periods that requires shifting.

@@

Standard seasonal plots break the overall time into two components: seasonal periods on the x-axis, and grouped by their corresponding lower-resolution time. For example, monthly data can be decomposed into months separated by years, and hourly data into hours grouped by days. Data collected at lower-level resolutions often exhibits more than one seasonal patterns. To discover typical seasonal or non-typical profiles, it is helpful to quickly browse through many possible periods. Interactivity ought to be enabled.

>>>>

The city of Melbourne has sensors installed to count hourly tallies of pedestrians in order to capture downtown daily rhythms (cite). Figure \ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at four locations, with the depiction of three pronounced slices in time. Figure \ref{fig:wrap-ped-1} unfolds all counts from January to May on their absolute time scales, facetted by four sensors. On March 16, Melbourne went to the stage three lockdown due to COVID-19, seeing a significant decline in traffic volume across the city. These lines are then folded into daily and weekly sections, shown as Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} respectively. Seasonal variations have been popped out to viewers, complementing the not-just-magnitude-drop story. The pre-lockdown period is coloured with dark green and lockdown with orange.

The wrapping procedure involves slicing time indices into seasonal periods of interest and their corresponding time dices. For example, hourly pedestrian data can be decomposed into 24-hour blocks grouped by all respective days, like Figure \ref{fig:wrap-ped-2}. Figure \ref{fig:wrap-ped} suggests that there could be more than one eye-catching slices out of many possible combinations, and thus repeated wrappings can be unwieldy. To visually locate an interesting slice, the \CRANpkg{tsibbletalk} package implements a shiny module, a pair of UI and server functions, to automate this wrapping procedure.

This shiny module, decoupled to `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece in a shiny application. Like all shiny modules, the first argument in both functions requires a user-supplied session id that must be unique. The UI function `tsibbleDiceUI()` simply shows a slider that animates or controls the number of periods to be diced. The workhorse is certainly the server function `tsibbleDiceServer()`, encapsulating the core algorithm that transforms data and sends messages to update the plot accordingly. It expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots). As the function name suggests, a (shared) tsibble is needed to start the engine. The `period` argument is to specify the minimum number of seasonal blocks, for example data shifted by "1 day", "2 days", or "1 week", etc. The following code chunk generates Figure \ref{fig:wrap-ped}.

@@

The initial state is

### 2020-09-06 tsibbletalk.tex
explained wrapping and shiny module functions

The city of Melbourne has sensors installed to count hourly tallies of
pedestrians in order to capture downtown daily rhythms (cite). Figure
\ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at
four locations, with the depiction of three pronounced slices in time.
Figure \ref{fig:wrap-ped-1} unfolds all counts from January to May on
their absolute time scales, facetted by four sensors. On March 16,
Melbourne went to the stage three lockdown due to COVID-19, seeing a
significant decline in traffic volume across the city. These lines are
then folded into daily and weekly sections, shown as Figure
\ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} respectively. Seasonal
variations have been popped out to viewers, complementing the
not-just-magnitude-drop data story. The pre-lockdown period is coloured
with dark green and lockdown with orange.

Figure \ref{fig:wrap-ped} suggests that there could be more than one
interesting time slices. The \CRANpkg{tsibbletalk} package provides a
pair of UI and server functions, as a shiny module, to help with finding
interesting seasonal slices in a shiny application. The pair, decoupled
to \texttt{tsibbleDiceUI()} and \texttt{tsibbleDiceServer()}, presents a
clean interface and forms a resusable piece. Like all shiny modules,
users should supply a unique session id. The UI function
\texttt{tsibbleDiceUI()} shows a slider that controls the number of
periods, and a plot specified by users. The server function
\texttt{tsibbleDiceServer()} is the workhorse, transforming data and
updating the plot. It expects a \texttt{ggplot} (converted to
\texttt{plotly} via \texttt{ggplotly()}) or \texttt{plotly} object. This
plot can be line charts, or other graphical elements (such as boxplots).
But it assumes that tsibble's time index is plotted on the x-axis. The
other mandatory argument is to specify the number of seasonal periods
that requires shifting.

@@

ui <- fluidPage(tsibbleDiceUI("dice"))
server <- function(input, output, session) {
  tsibbleDiceServer("dice", ggplotly(p_line, height = 700), period = "1 day")
}
shinyApp(ui, server)
\end{Sinput}
\end{Schunk}

Standard seasonal plots break the overall time into two components:
seasonal periods on the x-axis, and grouped by their corresponding
lower-resolution time. For example, monthly data can be decomposed into
months separated by years, and hourly data into hours grouped by days.
Data collected at lower-level resolutions often exhibits more than one
seasonal patterns. To discover typical seasonal or non-typical profiles,
it is helpful to quickly browse through many possible periods.
Interactivity ought to be enabled.

>>>>

The city of Melbourne has sensors installed to count hourly tallies of
pedestrians in order to capture downtown daily rhythms (cite). Figure
\ref{fig:wrap-ped} shows the first five months of 2020 foot traffic at
four locations, with the depiction of three pronounced slices in time.
Figure \ref{fig:wrap-ped-1} unfolds all counts from January to May on
their absolute time scales, facetted by four sensors. On March 16,
Melbourne went to the stage three lockdown due to COVID-19, seeing a
significant decline in traffic volume across the city. These lines are
then folded into daily and weekly sections, shown as Figure
\ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} respectively. Seasonal
variations have been popped out to viewers, complementing the
not-just-magnitude-drop story. The pre-lockdown period is coloured with
dark green and lockdown with orange.

The wrapping procedure involves slicing time indices into seasonal
periods of interest and their corresponding time dices. For example,
hourly pedestrian data can be decomposed into 24-hour blocks grouped by
all respective days, like Figure \ref{fig:wrap-ped-2}. Figure
\ref{fig:wrap-ped} suggests that there could be more than one
eye-catching slices out of many possible combinations, and thus repeated
wrappings can be unwieldy. To visually locate an interesting slice, the
\CRANpkg{tsibbletalk} package implements a shiny module, a pair of UI
and server functions, to automate this wrapping procedure.

This shiny module, decoupled to \texttt{tsibbleDiceUI()} and
\texttt{tsibbleDiceServer()}, presents a clean interface and forms a
resusable piece in a shiny application. Like all shiny modules, the
first argument in both functions requires a user-supplied session id
that must be unique. The UI function \texttt{tsibbleDiceUI()} simply
shows a slider that animates or controls the number of periods to be
diced. The workhorse is certainly the server function
\texttt{tsibbleDiceServer()}, encapsulating the core algorithm that
transforms data and sends messages to update the plot accordingly. It
expects a \texttt{ggplot} or \texttt{plotly} object, where one can plot
data using either lines or other graphical elements (such as boxplots).
As the function name suggests, a (shared) tsibble is needed to start the
engine. The \texttt{period} argument is to specify the minimum number of
seasonal blocks, for example data shifted by ``1 day'', ``2 days'', or
``1 week'', etc. The following code chunk generates Figure
\ref{fig:wrap-ped}.

@@

ui <- fluidPage(
  tsibbleDiceUI("dice")
)
server <- function(input, output, session) {
  tsibbleDiceServer("dice", ggplotly(p_line, height = 700), period = "1 day")
}
shinyApp(ui, server)
\end{Sinput}
\end{Schunk}

The initial state is
